valr: Reproducible Genome Interval Analysis in R

### 2017-06-20 inst/paper/figs.tex
fixed quotes and a few minor changes

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(valr)}
\KeywordTok{library}\NormalTok{(tidyverse)}
\KeywordTok{library}\NormalTok{(cowplot)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# `valr_example()` identifies the path of example files}
\NormalTok{bedfile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'genes.hg19.chr22.bed.gz'}\NormalTok{)}
\NormalTok{genomefile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{)}
\NormalTok{bgfile  <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hela.h3k4.chip.bg.gz'}\NormalTok{)}

@@

\NormalTok{p <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(res, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \NormalTok{.win_id, }\DataTypeTok{y =} \NormalTok{win_mean)) +}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{size =} \FloatTok{0.25}\NormalTok{) +}\StringTok{ }\KeywordTok{geom_pointrange}\NormalTok{(sd_limits, }\DataTypeTok{size =} \FloatTok{0.1}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{labels =} \NormalTok{x_labels, }\DataTypeTok{breaks =} \NormalTok{x_breaks) +}\StringTok{ }
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Position (bp from TSS)'}\NormalTok{) +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Signal'}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{() +}
\StringTok{  }\NormalTok{pub_theme}

\KeywordTok{plot_grid}\NormalTok{(p, }\DataTypeTok{align =} \StringTok{'h'}\NormalTok{, }\DataTypeTok{nrow =} \DecValTok{1}\NormalTok{, }\DataTypeTok{labels=}\StringTok{"AUTO"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(}\StringTok{"figure2.pdf"}\NormalTok{, }\DataTypeTok{width =} \FloatTok{5.6}\NormalTok{, }\DataTypeTok{height =} \DecValTok{4}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\CommentTok{# Set-up test genome}
\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{))}
\KeywordTok{write_tsv}\NormalTok{(genome, }\StringTok{"genome.txt"}\NormalTok{, }\DataTypeTok{col_names=}\OtherTok{FALSE}\NormalTok{)}

@@

\NormalTok{plot_a <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(res, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\KeywordTok{reorder}\NormalTok{(expr, time), }\DataTypeTok{y =} \NormalTok{time)) +}
\StringTok{  }\KeywordTok{geom_boxplot}\NormalTok{(}\DataTypeTok{width =} \FloatTok{0.75}\NormalTok{, }\DataTypeTok{size =} \FloatTok{0.33}\NormalTok{, }\DataTypeTok{outlier.shape =} \OtherTok{NA}\NormalTok{, }\DataTypeTok{fill =} \NormalTok{RColorBrewer::}\KeywordTok{brewer.pal}\NormalTok{(}\DecValTok{3}\NormalTok{, }\StringTok{"Set1"}\NormalTok{)[}\DecValTok{2}\NormalTok{]) +}
\StringTok{  }\KeywordTok{coord_flip}\NormalTok{() +}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() +}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}
    \DataTypeTok{y=}\StringTok{'execution time (seconds)'}\NormalTok{,}
    \DataTypeTok{x=}\StringTok{''}\NormalTok{,}
    \DataTypeTok{subtitle=}\KeywordTok{paste0}\NormalTok{(}\KeywordTok{comma}\NormalTok{(n), }\StringTok{" random x/y intervals,}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\KeywordTok{comma}\NormalTok{(nrep), }\StringTok{" repetitions"}\NormalTok{)) +}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{() +}
\StringTok{  }\NormalTok{pub_theme +}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}
    \DataTypeTok{legend.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{14}\NormalTok{),}
    \DataTypeTok{legend.title =} \KeywordTok{element_blank}\NormalTok{(),}
    \DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{,}
    \DataTypeTok{plot.subtitle =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{18}\NormalTok{),}
    \DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{)}
  \NormalTok{)}

\NormalTok{plot_b <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(dat, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\KeywordTok{reorder}\NormalTok{(func, time), }\DataTypeTok{y =} \NormalTok{time, }\DataTypeTok{fill =} \NormalTok{package)) +}
\StringTok{  }\KeywordTok{geom_boxplot}\NormalTok{(}\DataTypeTok{width =} \FloatTok{0.75}\NormalTok{, }\DataTypeTok{size =} \FloatTok{0.33}\NormalTok{, }\DataTypeTok{outlier.shape =} \OtherTok{NA}\NormalTok{) +}
\StringTok{  }\KeywordTok{scale_fill_brewer}\NormalTok{(}\DataTypeTok{palette =} \StringTok{"Set1"}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{coord_flip}\NormalTok{() +}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() +}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}
    \DataTypeTok{y=}\StringTok{'execution time (seconds)'}\NormalTok{,}
    \DataTypeTok{x=}\StringTok{''}\NormalTok{) +}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{() +}
\StringTok{  }\NormalTok{pub_theme +}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}
    \DataTypeTok{legend.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{18}\NormalTok{),}
    \DataTypeTok{legend.title =} \KeywordTok{element_blank}\NormalTok{(),}
    \DataTypeTok{legend.position =} \StringTok{"top"}\NormalTok{,}
    \DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{)}
  \NormalTok{)}

@@

@@

Using mysql

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(dplyr, }\DataTypeTok{warn.conflicts =} \NormalTok{F)}
\NormalTok{hg19 <-}\StringTok{ }\KeywordTok{db_ucsc}\NormalTok{(}\StringTok{"hg19"}\NormalTok{)}
\KeywordTok{tbl}\NormalTok{(hg19, }\StringTok{"refGene"}\NormalTok{)}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 0 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 4 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 5 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 6 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 7 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 8 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 0 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 4 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 5 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 6 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 7 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 8 imported}
\CommentTok{#> as numeric}
\CommentTok{#> # Source:   table<refGene> [?? x 16]}
\CommentTok{#> # Database: mysql 5.6.26-log [genomep@genome-mysql.cse.ucsc.edu:/hg19]}
\CommentTok{#>      bin         name chrom strand   txStart     txEnd  cdsStart    cdsEnd}
\CommentTok{#>    <dbl>        <chr> <chr>  <chr>     <dbl>     <dbl>     <dbl>     <dbl>}
\CommentTok{#>  1   585    NR_024540  chr1      -     14361     29370     29370     29370}
\CommentTok{#>  2  1759    NR_110022  chrX      + 153991016 154005964 154005964 154005964}
\CommentTok{#>  3   604    NR_033380  chrX      +   2527305   2575270   2575270   2575270}
\CommentTok{#>  4    75    NR_033381  chrY      +   2477305   2506370   2506370   2506370}
\CommentTok{#>  5   915    NM_182508 chr13      +  43355685  43365685  43358203  43362926}
\CommentTok{#>  6   923    NM_181845 chr19      +  44331472  44353050  44337656  44352793}
\CommentTok{#>  7   586    NR_039983  chr1      -    134772    140566    140566    140566}
\CommentTok{#>  8  1964    NR_028327  chr5      + 180750506 180755196 180755196 180755196}
\CommentTok{#>  9   126 NM_001271875 chr17      -  55938603  56032684  55943837  55962925}
\CommentTok{#> 10  2278    NR_125989  chr1      - 222001007 222014008 222014008 222014008}
\CommentTok{#> # ... with more rows, and 8 more variables: exonCount <dbl>,}
\CommentTok{#> #   exonStarts <chr>, exonEnds <chr>, score <int>, name2 <chr>,}
\CommentTok{#> #   cdsStartStat <chr>, cdsEndStat <chr>, exonFrames <chr>}
\end{Highlighting}
\end{Shaded}

>>>>

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(valr)}
\KeywordTok{library}\NormalTok{(tidyverse)}
\CommentTok{#> Loading tidyverse: ggplot2}
\CommentTok{#> Loading tidyverse: tibble}
\CommentTok{#> Loading tidyverse: tidyr}
\CommentTok{#> Loading tidyverse: readr}
\CommentTok{#> Loading tidyverse: purrr}
\CommentTok{#> Loading tidyverse: dplyr}
\CommentTok{#> Conflicts with tidy packages ----------------------------------------------}
\CommentTok{#> filter(): dplyr, stats}
\CommentTok{#> lag():    dplyr, stats}
\KeywordTok{library}\NormalTok{(cowplot)}
\CommentTok{#> }
\CommentTok{#> Attaching package: 'cowplot'}
\CommentTok{#> The following object is masked from 'package:ggplot2':}
\CommentTok{#> }
\CommentTok{#>     ggsave}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{bedfile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"genes.hg19.chr22.bed.gz"}\NormalTok{)}
\NormalTok{genomefile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.chrom.sizes.gz"}\NormalTok{)}
\NormalTok{bgfile  <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hela.h3k4.chip.bg.gz"}\NormalTok{)}

@@

\NormalTok{p <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(res, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \NormalTok{.win_id, }\DataTypeTok{y =} \NormalTok{win_mean)) +}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{size =} \FloatTok{0.25}\NormalTok{) +}\StringTok{ }\KeywordTok{geom_pointrange}\NormalTok{(sd_limits, }\DataTypeTok{size =} \FloatTok{0.1}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{labels =} \NormalTok{x_labels, }\DataTypeTok{breaks =} \NormalTok{x_breaks) +}\StringTok{ }
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Position (bp from TSS)"}\NormalTok{) +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{"Signal"}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{() +}
\StringTok{  }\NormalTok{pub_theme +}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{))}

\KeywordTok{plot_grid}\NormalTok{(p, }\DataTypeTok{align =} \StringTok{"h"}\NormalTok{, }\DataTypeTok{nrow =} \DecValTok{1}\NormalTok{, }\DataTypeTok{labels=}\StringTok{"AUTO"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(}\StringTok{"figure2.pdf"}\NormalTok{, }\DataTypeTok{width =} \FloatTok{5.6}\NormalTok{, }\DataTypeTok{height =} \FloatTok{5.6}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\CommentTok{# Set-up test genome}
\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.chrom.sizes.gz"}\NormalTok{))}
\KeywordTok{write_tsv}\NormalTok{(genome, }\StringTok{"genome.txt"}\NormalTok{, }\DataTypeTok{col_names=}\OtherTok{FALSE}\NormalTok{)}

@@

\NormalTok{plot_a <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(res, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\KeywordTok{reorder}\NormalTok{(expr, time), }\DataTypeTok{y =} \NormalTok{time)) +}
\StringTok{  }\KeywordTok{geom_boxplot}\NormalTok{(}\DataTypeTok{width =} \FloatTok{0.75}\NormalTok{, }\DataTypeTok{size =} \FloatTok{0.33}\NormalTok{, }\DataTypeTok{outlier.shape =} \OtherTok{NA}\NormalTok{, }\DataTypeTok{fill =} \NormalTok{grDevices::}\KeywordTok{grey.colors}\NormalTok{(}\DecValTok{2}\NormalTok{)[}\DecValTok{2}\NormalTok{]) +}
\StringTok{  }\KeywordTok{coord_flip}\NormalTok{() +}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() +}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}
    \DataTypeTok{y=}\StringTok{'execution time (seconds)'}\NormalTok{,}
    \DataTypeTok{x=}\StringTok{''}\NormalTok{,}
    \DataTypeTok{subtitle=}\KeywordTok{paste0}\NormalTok{(}\KeywordTok{comma}\NormalTok{(n), }\StringTok{" random x/y intervals,}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\KeywordTok{comma}\NormalTok{(nrep), }\StringTok{" repetitions"}\NormalTok{)) +}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{() +}
\StringTok{  }\NormalTok{pub_theme +}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}
    \DataTypeTok{legend.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{14}\NormalTok{),}
    \DataTypeTok{legend.title =} \KeywordTok{element_blank}\NormalTok{(),}
    \DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{,}
    \DataTypeTok{plot.subtitle =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{18}\NormalTok{),}
    \DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{)}
  \NormalTok{)}

\NormalTok{plot_b <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(dat, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\KeywordTok{reorder}\NormalTok{(func, time), }\DataTypeTok{y =} \NormalTok{time, }\DataTypeTok{fill =} \NormalTok{package)) +}
\StringTok{  }\KeywordTok{geom_boxplot}\NormalTok{(}\DataTypeTok{width =} \FloatTok{0.75}\NormalTok{, }\DataTypeTok{size =} \FloatTok{0.33}\NormalTok{, }\DataTypeTok{outlier.shape =} \OtherTok{NA}\NormalTok{) +}
\StringTok{  }\KeywordTok{scale_fill_grey}\NormalTok{() +}\StringTok{ }
\StringTok{  }\KeywordTok{coord_flip}\NormalTok{() +}
\StringTok{  }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =} \KeywordTok{seq_along}\NormalTok{(}\KeywordTok{unique}\NormalTok{(dat$func)) +}\StringTok{ }\FloatTok{0.5}\NormalTok{,}
             \DataTypeTok{color =} \StringTok{"grey"}\NormalTok{) +}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() +}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}
    \DataTypeTok{y=}\StringTok{'execution time (seconds)'}\NormalTok{,}
    \DataTypeTok{x=}\StringTok{''}\NormalTok{) +}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{() +}
\StringTok{  }\NormalTok{pub_theme +}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}
    \DataTypeTok{legend.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{18}\NormalTok{),}
    \DataTypeTok{legend.title =} \KeywordTok{element_blank}\NormalTok{(),}
    \DataTypeTok{legend.position =} \StringTok{"top"}\NormalTok{,}
    \DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{)}
  \NormalTok{)}

@@

\subsubsection{file I/O}\label{file-io}

@@

\subsubsection{syntax demo}\label{syntax-demo}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(valr)}
\KeywordTok{library}\NormalTok{(dplyr)}

\NormalTok{snps <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.snps147.chr22.bed.gz"}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{"genes.hg19.chr22.bed.gz"}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}

\CommentTok{# find snps in intergenic regions}
\NormalTok{intergenic <-}\StringTok{ }\KeywordTok{bed_subtract}\NormalTok{(snps, genes)}
\CommentTok{# distance from intergenic snps to nearest gene}
\NormalTok{nearby <-}\StringTok{ }\KeywordTok{bed_closest}\NormalTok{(intergenic, genes)}

\NormalTok{nearby %>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\KeywordTok{starts_with}\NormalTok{(}\StringTok{"name"}\NormalTok{), .overlap, .dist) %>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\KeywordTok{abs}\NormalTok{(.dist) <}\StringTok{ }\DecValTok{1000}\NormalTok{)}
\CommentTok{#> # A tibble: 285 x 4}
\CommentTok{#>         name.x            name.y .overlap .dist}
\CommentTok{#>          <chr>             <chr>    <int> <int>}
\CommentTok{#>  1   rs2261631             P704P        0  -267}
\CommentTok{#>  2 rs570770556             POTEH        0  -912}
\CommentTok{#>  3 rs538163832             POTEH        0  -952}
\CommentTok{#>  4   rs9606135            TPTEP1        0  -421}
\CommentTok{#>  5  rs11912392 ANKRD62P1-PARP4P3        0   104}
\CommentTok{#>  6   rs8136454          BC038197        0   355}
\CommentTok{#>  7   rs5992556              XKR3        0  -455}
\CommentTok{#>  8 rs114101676              GAB4        0   473}
\CommentTok{#>  9  rs62236167             CECR7        0   261}
\CommentTok{#> 10   rs5747023             CECR1        0  -386}
\CommentTok{#> # ... with 275 more rows}
\end{Highlighting}
\end{Shaded}

\subsubsection{Figure 1 code demo}\label{figure-1-code-demo}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end,}
  \StringTok{"chr1"}\NormalTok{, }\DecValTok{25}\NormalTok{,     }\DecValTok{50}\NormalTok{,}
  \StringTok{"chr1"}\NormalTok{, }\DecValTok{100}\NormalTok{,    }\DecValTok{125}
\NormalTok{)}

\NormalTok{y <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end,}
  \StringTok{"chr1"}\NormalTok{, }\DecValTok{30}\NormalTok{,     }\DecValTok{75}
\NormalTok{)}

\KeywordTok{bed_glyph}\NormalTok{(}\KeywordTok{bed_intersect}\NormalTok{(x, y))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end,}
  \StringTok{"chr1"}\NormalTok{,      }\DecValTok{1}\NormalTok{,      }\DecValTok{50}\NormalTok{,}
  \StringTok{"chr1"}\NormalTok{,      }\DecValTok{10}\NormalTok{,     }\DecValTok{75}\NormalTok{,}
  \StringTok{"chr1"}\NormalTok{,      }\DecValTok{100}\NormalTok{,    }\DecValTok{120}
\NormalTok{)}

\KeywordTok{bed_glyph}\NormalTok{(}\KeywordTok{bed_merge}\NormalTok{(x))}
\end{Highlighting}
\end{Shaded}

\subsubsection{grouping data demo}\label{grouping-data-demo}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end, ~strand,}
  \StringTok{"chr1"}\NormalTok{, }\DecValTok{1}\NormalTok{,      }\DecValTok{100}\NormalTok{,  }\StringTok{"+"}\NormalTok{,}
  \StringTok{"chr1"}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{"+"}\NormalTok{,}
  \StringTok{"chr2"}\NormalTok{, }\DecValTok{100}\NormalTok{,    }\DecValTok{200}\NormalTok{,  }\StringTok{"-"}
\NormalTok{)}

\NormalTok{y <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end, ~strand,}
  \StringTok{"chr1"}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{125}\NormalTok{,  }\StringTok{"+"}\NormalTok{,}
  \StringTok{"chr1"}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{"-"}\NormalTok{,}
  \StringTok{"chr2"}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{"+"}
\NormalTok{)}

\CommentTok{# intersect tbls by strand}
\NormalTok{x <-}\StringTok{ }\KeywordTok{group_by}\NormalTok{(x, strand)}
\NormalTok{y <-}\StringTok{ }\KeywordTok{group_by}\NormalTok{(y, strand)}

\KeywordTok{bed_intersect}\NormalTok{(x, y)}
\CommentTok{#> # A tibble: 2 x 8}
\CommentTok{#>   chrom start.x end.x strand.x start.y end.y strand.y .overlap}
\CommentTok{#>   <chr>   <dbl> <dbl>    <chr>   <dbl> <dbl>    <chr>    <int>}
\CommentTok{#> 1  chr1       1   100        +      50   125        +       50}
\CommentTok{#> 2  chr1      50   150        +      50   125        +       75}
\end{Highlighting}
\end{Shaded}

Comparisons between intervals on opposite strands are done using the
\texttt{flip\_strands()} function:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x <-}\StringTok{ }\KeywordTok{group_by}\NormalTok{(x, strand)}

\NormalTok{y <-}\StringTok{ }\KeywordTok{flip_strands}\NormalTok{(y)}
\NormalTok{y <-}\StringTok{ }\KeywordTok{group_by}\NormalTok{(y, strand)}

\KeywordTok{bed_intersect}\NormalTok{(x, y)}
\CommentTok{#> # A tibble: 3 x 8}
\CommentTok{#>   chrom start.x end.x strand.x start.y end.y strand.y .overlap}
\CommentTok{#>   <chr>   <dbl> <dbl>    <chr>   <dbl> <dbl>    <chr>    <int>}
\CommentTok{#> 1  chr2     100   200        -      50   150        -       50}
\CommentTok{#> 2  chr1       1   100        +      50   150        +       50}
\CommentTok{#> 3  chr1      50   150        +      50   150        +      100}
\end{Highlighting}
\end{Shaded}

\subsubsection{Figure 2 code demo}\label{figure-2-code-demo}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{bedfile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"genes.hg19.chr22.bed.gz"}\NormalTok{)}
\NormalTok{genomefile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.chrom.sizes.gz"}\NormalTok{)}
\NormalTok{bgfile  <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hela.h3k4.chip.bg.gz"}\NormalTok{)}

\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(bedfile, }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(genomefile)}
\NormalTok{y <-}\StringTok{ }\KeywordTok{read_bedgraph}\NormalTok{(bgfile)}
\end{Highlighting}
\end{Shaded}

Then we generate 1 bp intervals to represent transcription start sites
(TSSs). We focus on \texttt{+} strand genes, but \texttt{-} genes are
easily accomodated by filtering them and using
\texttt{bed\_makewindows()} with \texttt{reversed} window numbers.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# generate 1 bp TSS intervals, "+" strand only}
\NormalTok{tss <-}\StringTok{ }\NormalTok{genes %>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(strand ==}\StringTok{ "+"}\NormalTok{) %>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{end =} \NormalTok{start +}\StringTok{ }\DecValTok{1}\NormalTok{)}

\NormalTok{x}
\CommentTok{#> # A tibble: 13,530 x 7}
\CommentTok{#>    chrom    start      end      name score strand .win_id}
\CommentTok{#>    <chr>    <int>    <int>     <chr> <chr>  <chr>   <int>}
\CommentTok{#>  1 chr22 16161065 16161115 LINC00516     3      +       1}
\CommentTok{#>  2 chr22 16161115 16161165 LINC00516     3      +       2}
\CommentTok{#>  3 chr22 16161165 16161215 LINC00516     3      +       3}
\CommentTok{#>  4 chr22 16161215 16161265 LINC00516     3      +       4}
\CommentTok{#>  5 chr22 16161265 16161315 LINC00516     3      +       5}
\CommentTok{#>  6 chr22 16161315 16161365 LINC00516     3      +       6}
\CommentTok{#>  7 chr22 16161365 16161415 LINC00516     3      +       7}
\CommentTok{#>  8 chr22 16161415 16161465 LINC00516     3      +       8}
\CommentTok{#>  9 chr22 16161465 16161515 LINC00516     3      +       9}
\CommentTok{#> 10 chr22 16161515 16161565 LINC00516     3      +      10}
\CommentTok{#> # ... with 13,520 more rows}
\end{Highlighting}
\end{Shaded}

Now we use the \texttt{.win\_id} group with \texttt{bed\_map()} to
caluclate a sum by mapping \texttt{y} signals onto the intervals in
\texttt{x}. These data are regrouped by \texttt{.win\_id} and a summary
with \texttt{mean} and \texttt{sd} values is calculated.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# map signals to TSS regions and calculate summary statistics.}
\NormalTok{res <-}\StringTok{ }\KeywordTok{bed_map}\NormalTok{(x, y, }\DataTypeTok{win_sum =} \KeywordTok{sum}\NormalTok{(value, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{)) %>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(.win_id) %>%}
\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{win_mean =} \KeywordTok{mean}\NormalTok{(win_sum, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \DataTypeTok{win_sd =} \KeywordTok{sd}\NormalTok{(win_sum, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{))}

\NormalTok{res}
\CommentTok{#> # A tibble: 41 x 3}
\CommentTok{#>    .win_id win_mean    win_sd}
\CommentTok{#>      <int>    <dbl>     <dbl>}
\CommentTok{#>  1       1 100.8974  85.83423}
\CommentTok{#>  2       2 110.6829  81.13521}
\CommentTok{#>  3       3 122.9070  99.09635}
\CommentTok{#>  4       4 116.2800  96.30098}
\CommentTok{#>  5       5 116.3500 102.33773}
\CommentTok{#>  6       6 124.9048  95.08887}
\CommentTok{#>  7       7 122.9437  94.39792}
\CommentTok{#>  8       8 127.5946  91.47407}
\CommentTok{#>  9       9 130.2051  95.71309}
\CommentTok{#> 10      10 130.1220  88.82809}
\CommentTok{#> # ... with 31 more rows}
\end{Highlighting}
\end{Shaded}

Finally, these summary statistics are used to construct a plot that
illustrates histone density surrounding TSSs.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(ggplot2)}
